Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-04 paper.tex
Edits

Overall, the data is quite sparse. As mentioned before, the experts were allowed to update their forecasts as long as the question remained open. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day. Therefore a typical expert made a forecast about every 62 days, resulting to an average response rate of around 13.5 forecasts per day from a very large and diverse group of experts. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Even though it is unlikely that any single expert has all the available information, a large and diverse group of experts may share a considerable portion of this information. The so called \textit{wisdom of the crowd} is therefore an attractive proxy for the amount of information available at any given time point.

If a model is to efficiently use this proxy for estimating the hidden process $X_{t,k}$, it must be able to discriminate between noise and signal as experts may believe in false information, hide their true beliefs on an ethical basis, or be biased for many other reasons. Since in our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief, an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 Similarly to the observed process, the hidden process  has a systematic and a random component. The \textit{random fluctuations} are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$, on the other hand, allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

>>>>

The experts made updates on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days. The average group-level response rate was around 13.5 forecasts per day. Given that the group of experts is large and diverse, the data is very sparse. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that this probability forecast 
%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Any single expert is unlikely to have enough information to consistently report $\logit^{-1}(X_{t,k})$. He may also believe in false information, hide his true belief for ethical bases, or be biased for many other reasons. A large and diverse group of experts, on the other hand, may share a considerable portion of the available information. But if they are not collaborating and sharing information efficiently, their beliefs should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.

%A model should therefore aim to take into account the collective opinion of a group of experts. 
%
%By taking into account the collective opinion the group of experts may be able to 
%
%
%. The collective wisdom of this group is therefore an attractive proxy for the amount of information available at any given time point. 
%
%The experts, however, may not be collaborating and sharing information efficiently with each other.

%Therefore, if a model is to efficiently use the collective wisdom for estimating $X_{t,k}$, it must be able to discriminate between noise and signal.

 In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$, on the other hand, allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.
